---
title: "Investigating Exponential Distributions and CLT"
author: "Tiago Vinícius"
date: "06/14/2015"
output: pdf_document
---
<br><br>

#Overview

It will be shown comparisons about Exponential distribution and Central Limit Theorem (CLT). Investigation is about assumption that under a large number of simulations (i.e. Law of Large Numbers), Exponential distribution converges to Gaussian distribution. The theorical mean and standard deviation are 1/lambda, so I will compare...
<br>

#Simulations

First, let's simulate an exponential distribution with n = 40 and lambda = 0.2 and repeat simulation in 1000 times:

```{r, echo=T}
#Reproducible seed
set.seed(17)

lambda <- 0.2
n = 40
rep <- 1000

#exponential distribution replicated 1000 times
exp <- replicate(rep, rexp(n, lambda))

#Resultant mean
means_exp <- apply(exp,2,mean)

```

#Sample Mean versus Theoretical Mean

1. Show the sample mean and compare it to the theoretical mean of the distribution.
<br>
After simulation, we compare sample mean and theoretical mean:
```{r}
sample_mean <- mean(means_exp)
sample_mean

theoretical_mean <- 1/lambda
theoretical_mean
```

#Sample Variance versus Theoretical Variance
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
```{r}
sample_variance <- var(means_exp)
sample_variance

theoretical_variance <- (1/lambda)^2
theoretical_variance
```

#Distribution

3. Show that the distribution is approximately normal.

#Summary