---
title: "Motor Trend Regression Analysis"
author: "Tiago V R da Costa"
date: "20/06/2015"
output: pdf_document
---
\vspace*{1\baselineskip}
  
#Questions
Exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

- "Is an automatic or manual transmission better for MPG?"
- "Quantify the MPG difference between automatic and manual transmissions"

\vspace*{1\baselineskip}

#Executive Summary
In this paper, to achieve our goal by answering these questions, we perform exploratory analysis to visualize *mpg* and *am* variables relations. In a second step we perform hyphotesis test in order to evaluate if manual transmission is better than automatic transmission for mpg performance metric.
\vspace*{1\baselineskip}
In the end, we perform multiple linear regression analysis, using backward approach, over MPG as outcome and all variables at first, to select best model variables. After that, we will be able to quantify *mpg* difference between *am* variable possible values.
\vspace*{1\baselineskip}


```{r}
#load dataset
data(mtcars)

#sample
head(mtcars)

#Boxplot automatic vs manual
boxplot(mpg ~ am, data = mtcars, col="red", 
        xlab = "transmission", ylab = "miles per gallon", main = "Automatic vs manual")
```

\vspace*{1\baselineskip}Boxplot shows some trend that automatic tramission has less mpg performance than manual transmission. That is what we want to prove in next section.

\vspace*{2\baselineskip}

#Is an automatic or manual transmission better for MPG?

Our null hyphotesis stands that manual transmission has better performance for mpg, because of previous exploratory bloxplot analysis. Hyphotesis test can show if this is a true assumption.

```{r}
#Split group
t0 <- subset(mtcars, am=="0") #automatic
t1 <- subset(mtcars, am=="1") #manual

t.test(t0,t1, paired = FALSE, var.equal = TRUE)
```
*Answer: As p-value > 0.05, we do not have enough arguments to reject null hyphotesis, therefore, manual transmission actually is better for mpg.*

\vspace*{1\baselineskip}

#Quantify the MPG difference between automatic and manual transmissions

This time, we select all variables to perform base prediction of outcome variable *mpg*. In order to automatically select best model variables, we use *backward step* variable selection.


```{r}
#model considering all variables
fit <- lm(mpg ~ ., data = mtcars)

#backward strategy: all variables are considered, 
#than model excludes less discrimant variables 
backward <- step(fit, direction = "backward")
summary(backward)
```
After, step backward select the *lm(formula = mpg ~ wt + qsec + am, data = mtcars)* as best model. Considering this model, difference between automatic and manual transmissions is given by *am* estimated value:
```{r}
summary(backward)$coef
```

\vspace*{1\baselineskip}

*Answer: the MPG difference between automatic and manual transmission is about 2.9358 miles per gallon.*